---
title: "Statistics: Analysis of Variance (ANOVA)"
author: "Scott Schumacker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***Loading Packages***
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(tidyr)
```

## Creating a data set
For this example of an analysis of variance, we can imagine 3 blood pressure drugs to see which drug has a statistical difference in lowering blood pressure.

We can create three data sets, one for each example drug. We can use the rnorm function to create normal distributions.
```{r}
DrugA <- rnorm(100, mean = 80, sd = 1)
DrugB <- rnorm(100, mean = 95, sd = 1)
DrugC <- rnorm(100, mean = 83, sd = 1)

df <- data.frame(DrugA, DrugB, DrugC)

df2 <- df %>% pivot_longer(
  cols = `DrugA`:`DrugC`,
  names_to = "Drug",
  values_to = "Blood_Pressure"
)
head(df2)
```

After obtaining our data set we can visually inspect the distributions to better understand the shape of the data.

### Checking for normality
```{r}
ggplot(df2, aes(x = Blood_Pressure, fill = Drug)) +
  geom_histogram(color = "black", alpha = 0.6) +
  theme_classic() +
  xlab("Diastolic Blood Pressure")
```

***Checking for normality***

### Statistical Analysis

***ANOVA***

### Results
