---
title: "Statistics: Analysis of Variance (ANOVA)"
author: "Scott Schumacker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project will show an example Analysis of Variance (ANOVA) on 3 different example blood pressure medications to see if there is a statistically significant difference in lowering blood pressure.

***Loading Packages***
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(tidyr)
```

## Creating a data set
For this example of an analysis of variance, we can imagine 3 different blood pressure medications to see which drug has a statistically significant difference in lowering blood pressure.

We can create three data sets, one for each example medication. We can use the rnorm function to create normal distributions. In addition, we can keep things simple and keep the sample size for each group at 100 samples.
```{r}
DrugA <- rnorm(100, mean = 80, sd = 1)
DrugB <- rnorm(100, mean = 95, sd = 1)
DrugC <- rnorm(100, mean = 95, sd = 1)

df <- data.frame(DrugA, DrugB, DrugC)

df2 <- df %>% pivot_longer(
  cols = `DrugA`:`DrugC`,
  names_to = "Drug",
  values_to = "Blood_Pressure"
)
head(df2)
```

After obtaining our data set, we can visually inspect the distributions to better understand the shape of the data.

```{r}
ggplot(df2, aes(x = Blood_Pressure, fill = Drug)) +
  geom_histogram(color = "black", alpha = 0.6) +
  theme_classic() +
  xlab("Diastolic Blood Pressure")
```

After visualizing the data, we can see that all three distributions follow a normal distribution pattern. In addition, we can see that there may be a difference between Medication A and Medications B and C. 

We can confirm that the three groups follow a normal distribution by testing for normality by running the Shaprio-Wilk Test with an alpha = 0.05.

### Checking for normality and homoscedasticity

***Checking for normality***
```{r}
shapiro.test(DrugA)
shapiro.test(DrugB)
shapiro.test(DrugC)
```
Because our p-value is above our alpha of 0.05, we can assume that the data is normally distributed.

Now that we know that the three sample groups follow a normal distribution, we can check the assumption of homogeneity of variance (homoscedasticity) before running our t-test. To test homogeneity of variance, we can run Bartlett's Test.

We can run Bartlett's Test over Fisher's F Test because Bartlett's Test can accomodate more than two groups. We can run Bartlett's Test with an alpha of 0.05.

***Checking for homoscedasticity***
```{r}
bartlett.test(Blood_Pressure ~ Drug, data = df2)
```
Because our p-value is above our alpha of 0.05, we can assume homogeneity of variance between the three groups.

Now that we know that our data is randomly sampled, follows a normal distribution, is continuous, and there is homogeneity of variance between the three groups, we can run the ANOVA.

### Statistical Analysis

Before running the ANOVA we need to specify our hypothesis: <br>
H0: The difference in the means between the groups is 0 (null hypothesis) <br>
Ha: The difference in the means between the groups is not 0 (alternative hypothesis)

We can set our alpha to 0.05 for this example. This means that we are looking for a p-value below 0.05 to reject the null hypothesis.

***ANOVA***
```{r}
analysisOut <- aov(Blood_Pressure ~ Drug, data = df2)
summary(analysisOut)
```
### ANOVA Results

Our p-value is less than 0.01 which means that we can reject our null hypothesis. We can assume that there is a statistically significant difference between the three groups in terms of blood pressure.

### Tukey's Honest Significant Difference (HSD) Test

From our ANOVA we know that there is a significant difference among the three groups but we don't know where that difference is. In order to get a better idea of where the difference is, we can run Tukey's Honest Significant Difference test.
```{r}
TukeyHSD(analysisOut, conf.level = 0.95)
```
### Results 

We can see that there is a statistically significant difference between Drug B and Drug A and there is a statistically significant difference between Drug C and Drug A and no significant difference between Drug B and Drug C. This points us to the conclusion that Drug A is may be more optimal for lowering blood pressure.